---
title: "Milwaukee's Share of Low-Performing Schools"
author: "Spencer Schien"
date: "9/25/2019"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
```


```{r}
library(DBI)
library(tidyverse)

con <- dbConnect(RSQLite::SQLite(), "C:/Users/Spencer/OneDrive - Schools That Can Milwaukee/Data & Impact/R Files/schools_db/school_db.sqlite")

rc <- dbReadTable(con, "report_cards")

low_performers <- c("Alternate Rating - Needs Improvement",
                    "Fails to Meet Expectations",
                    "Fails to Meed Expectations^")

rc_18 <- rc %>%
  filter(school_year == "2017-18" & !(has_2_rc == 1 & report_card_type == "Private - Choice Students")) %>%
  mutate(mke_indicator = ifelse((city == "Milwaukee" & (locale_description == "City" | locale_description == "NA")), 1, 0),
         low_perf = ifelse(overall_rating %in% low_performers, 1, 0))


performance <- rc_18 %>%
  group_by(mke_indicator) %>%
  summarise(total = n(),
            low = sum(low_perf))
```

The purpose of this document is to answer the following question:  **What share of Wisconsin's low-performing schools are in Milwaukee?**

First, it is instructive to know what share of Wisconsin schools are in Milwaukee in general.

* **Milwaukee**: `r sum(rc_18$mke_indicator)` (`r paste(round(sum(rc_18$mke_indicator)/nrow(rc_18)*100, 1), "%", sep = "")`) 
* **Not Milwaukee**: `r sum(rc_18$mke_indicator == 0)` (`r paste(round(sum(rc_18$mke_indicator == 0)/nrow(rc_18)*100, 1), "%", sep = "")`)

### Performance as indicated by the School Report Card Overall Ratings

First, we will label schools as *low-performing* based on having the following School Report Card ratings:  
  
* Alternate Rating - Needs Improvemen
* Fails to Meet Expectation
* Fails to Meed Expectations^ 
  
By this definition, a Milwaukee school is 20% more likely to be low-performing than a school not in Milwaukee, and a school not in Milwaukee is 18% LESS likely to be low-performing.

```{r}
performance %>%
  mutate(mke_indicator = ifelse(mke_indicator == 1, "Milwaukee", "Not Milwaukee")) %>%
  select("Milwaukee Indicator" = mke_indicator,
         "Total Schools" = total,
         "Low Performing Schools" = low) %>%
  kable()
```

### Performance as indicated by Spencer's Benchmarking

Next, we will label schools as *low-performing* based on the the Benchmarking model Spencer created.  **It should be noted that the original model was optimized for Milwaukee, meaning no out-state schools were included.**  Therefore, the results listed below should be interpreted with caution.

Also recall that this model depends on the Achievement Score, and schools are omitted when they don't have that score.  This is why the number of *Total Schools* are different.

```{r}

library(broom)

nested <- rc_18 %>%
  nest(-grade_band)

mod_fun_clust <- function(df) {
  lm(sch_ach ~ per_ed + per_swd + per_b_aa, data = df)
}

nested_ach <- nested %>%
  mutate(model = map(data, mod_fun_clust),
         tidied = map(model, tidy),
         augmented = map2(model, data, augment_columns)) %>%
  arrange(grade_band)

modeled_ach <- nested_ach  %>%
  unnest(augmented) %>%
  mutate(bench_low_ind = ifelse(.resid >= mean(.resid, na.rm = TRUE) - 2 * sd(.resid, na.rm = TRUE), 0, 1))


bench_perf <- modeled_ach %>%
  group_by(mke_indicator) %>%
  summarise(total = n(),
            low = sum(bench_low_ind)) 

bench_perf %>%
  mutate(mke_indicator = ifelse(mke_indicator == 1, "Milwaukee", "Not Milwaukee")) %>%
  select("Milwaukee Indicator" = mke_indicator,
         "Total Schools" = total,
         "Low Performing Schools" = low) %>%
  kable()
```

This table labels as low-performing those schools whose *Difference from Benchmark* is greater than two standard deviations **below** the mean--put another way, it's about the bottom 2.5%.

By this definition, a Milwaukee School is 18% more likely to be low-performing than a school not in Milwaukee, but a school not in Milwaukee is only 2% less likely to be low-performing.